---
title: "Chapter 16 iteration"

subtitle: "Using purrr::map"

author: |
  | McEwen Khundi
  |
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    df_print: paged
    theme: spacelab
    toc: true
---

<br>

## 1. Backgound

<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, message=FALSE}
library(tidyverse)    #for data manipulation
library(pmthemes)     #for ggplot themes
library(knitr)        #for tables

```

<br>

## 3. Import datasets

Import data required for the analysis.

```{r import}

linelist <- read_rds(here::here("linelist_cleaned.rds"))

```

<br>

## 4. Some brief introduction on defining functions in R

-   Why do we define functions

    -   To reduce repetition

    -   Ease reuse of code

    -   Abstract out the details

-   For this exercise assume R does not have an inbuilt mean function.

    -   Find the mean age, generation, weight, ht_cm

```{r mean}

#mean age
linelist %>%
  pull(age_years) %>% ## get the sum 
  sum(na.rm = TRUE)/ ## devide
linelist %>%
  filter(!is.na(age_years)) %>% ## get the total non missing values
    nrow()

#mean generation
linelist %>%
  pull(generation) %>% ## get the sum 
  sum(na.rm = TRUE)/ ## devide
linelist %>%
  filter(!is.na(generation)) %>% ## get the total non missing values
    nrow()

#mean weight kg
linelist %>%
  pull(wt_kg) %>% ## get the sum 
  sum(na.rm = TRUE)/ ## devide
linelist %>%
  filter(!is.na(wt_kg)) %>% ## get the total non missing values
    nrow()


#mean height cm
linelist %>%
  pull(ht_cm) %>% ## get the sum 
  sum(na.rm = TRUE)/ ## devide
linelist %>%
  filter(!is.na(ht_cm)) %>% ## get the total non missing values
    nrow()

```

## 5. Define a function for mean (to reduce repetition)

-   And use function

```{r}

#tidyverse version
mean_fn_tidy <- function(num_var){
  linelist %>%
  pull({{num_var}}) %>% ## get the sum 
  sum(na.rm = TRUE)/ ## devide
linelist %>%
  filter(!is.na({{num_var}})) %>% ## get the total non missing values
    nrow()
}

#a baseR version
mean_fn_base <- function(num_var){
  sum(linelist[[num_var]], na.rm = TRUE)/
  sum(!is.na(linelist[[num_var]]))
}

mean_fn_base("age_years")

mean_fn_tidy("age_years")

mean_fn_tidy(generation)

mean_fn_tidy(wt_kg)

mean_fn_tidy(ht_cm)

```

## 5. How can we reduce repetition (iteration)

```{r iteration}

num_vec <- vector(mode="numeric", length = 2)

step = 1
for(i in c("age_years", "generation")){
  
    num_vec[step] <- mean_fn_tidy(i)
    
    step = step + 1
}















```

## 5. Disadvantages of a for loop

-   Set up the sequence

-   Set up the container

Introduce purrr::map

```{r maps}


map(c("age_years", "generation"), mean_fn_tidy)



```

## 5. Passing arguments to functions in map

-   Use the R based mean

-   It is generalised can work with any other dataset

```{r}



map(select(linelist,"age_years", "generation"), mean, na.rm = TRUE)



```

## 6. Anonymous functions in Maps

-   Introduce the ., .x and \~ notation

```{r}


map(select(linelist,"age_years", "generation"),
    
  function(x){
    
  sum(x, na.rm = TRUE)/
  sum(!is.na(x))
    
  })


map(select(linelist,"age_years", "generation"),
    
  ~sum(.x, na.rm = TRUE)/sum(!is.na(.x))
    
  )


```

## 6. References

-   <https://epirhandbook.com/en/iteration-loops-and-lists.html>

-   <https://r4ds.had.co.nz/iteration.html>

-   <https://www.youtube.com/watch?v=GxvccD8K49M>

<br>

## 7. Reproducibility

This reproduction of the analysis was run by:

```{r sysinfo, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

sysinfo <- Sys.info()

sysinfo <- data.frame(keyName=names(sysinfo), value=sysinfo, row.names=NULL)

sysinfo %>% kable()
```

Analysis was run at **`r Sys.time()`**, and using the following Session Info:

```{r sessioninfo, echo=FALSE, results='markdown', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```
