---
title: "R r4ds book club chapter 25"

subtitle: "Many Models"

author: |
  | McEwen Khundi
  |
  | London School of Hygiene & Tropical Medicine, London, UK
  | Malawi-Liverpool-Wellcome Clinical Research Programme, Blantyre, Malawi
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    df_print: paged
    theme: spacelab
    toc: true
---

<br>

## 1. Backgound

### Now that we are close to the end of the book we need to start thinking about the next book to cover below are my suggested books.

1.  <https://epirhandbook.com/index.html>

2.  <https://openintro-ims.netlify.app/>

<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, message=FALSE}
library(tidyverse)    #for data manipulation
library(modelr)
library(pmthemes)     #for ggplot themes
library(knitr)        #for tables

```

<br>

## 3 Main ideas about this model

1.  Using many simple models to better understand complex datasets

2.  Using list-columns to store arbitrary data structures in a data frame.

3.  Using the **broom** package, by David Robinson, to turn models into tidy data. (Formatting model output)

## 3. Data will be form gapminder

life expectancy around the world. It's a small dataset but it illustrates how important modelling can be for improving your visualisations.

This data was popularised by Hans Rosling, a Swedish doctor and statistician. <https://www.youtube.com/watch?v=jbkSRLYSojo>.

```{r import}

library(gapminder)
gapminder

```

<br>

## 4. Question for demo: "How does life expectancy (`lifeExp`) change over time (`year`) for each country (`country`)?"

This to note from the graph

1.  Overall trend inn life expectancy is improving

2.   Some countries are not doing so well

```{r tidy}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3)
```

<br>

## 5. What can we do to remove the time trend?

1.  Need to remove the effect of time by fitting a linear trend.

2.  The model will capture the steady growth so that the left patterns will be due to other factors and not time.

3.  First step is to do it for a single country they used New Zealand in the book

```{r removetrend}
mw <- filter(gapminder, country == "Malawi")
mw %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_line() + 
  ggtitle("Full data = ")

mw_mod <- lm(lifeExp ~ year, data = mw)
mw %>% 
  add_predictions(mw_mod) %>%
  ggplot(aes(year, pred)) + 
  geom_line() + 
  ggtitle("Linear trend + ")

mw %>% 
  add_residuals(mw_mod) %>% 
  ggplot(aes(year, resid)) + 
  geom_hline(yintercept = 0, colour = "white", size = 3) + 
  geom_line() + 
  ggtitle("Remaining pattern")
```

<br>

## 6. Repeat task for all countries

Using nested data frames

1.  Creates a dataframe that is one group per row with a data column

2.  Differs from a grouped dataset

```{r}
by_country <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()


by_country


by_country$data[[1]]
```

```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}

## We have two options of using the function above
models <- map(by_country$data, country_model)

models

by_country <- by_country %>% 
  mutate(model = map(data, country_model))

#Big idea is related obj are kept together
by_country

```

**Unnesting getting back the values in nested data column**

```{r}
by_country <- by_country %>% 
  mutate(
    resids = map2(data, model, add_residuals)
  )
by_country

resids <- unnest(by_country, resids)
resids

```

**Can plot the residuals to look at the patterns after removing time trend effect**

```{r}
resids %>% 
  ggplot(aes(year, resid)) +
    geom_line(aes(group = country), alpha = 1 / 3) + 
    geom_smooth(se = FALSE)

#By continent
resids %>% 
  ggplot(aes(year, resid, group = country)) +
    geom_line(alpha = 1 / 3) + 
    facet_wrap(~continent)
```

<br>

## 7. Model quality checks using functions from the broom package

```{r}
broom::glance(mw_mod)


#Easy to exted to the rest of countries

 glance <- by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)

```

## 8 Lets focus on the r.squared stats

```{r}

glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_jitter(width = 0.5)

```

### 9 Countries with worst model fits

```{r}
bad_fit <- filter(glance, r.squared < 0.25)

gapminder %>% 
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()

```

<br>

## 10. References

1.  <https://r4ds.had.co.nz/many-models.html#list-columns-1>

2.  I also haven't covered sections 25.4 to 25.6

<br>

## 11. Reproducibility

This reproduction of the analysis was run by:

```{r sysinfo, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

sysinfo <- Sys.info()

sysinfo <- data.frame(keyName=names(sysinfo), value=sysinfo, row.names=NULL)

sysinfo %>% kable()
```

Analysis was run at **`r Sys.time()`**, and using the following Session Info:

```{r sessioninfo, echo=FALSE, results='markdown', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```
