---
title: "R for data science book review: Data visualisation chapter 3"
author: "McEwen Khundi"
date: "11/12/2020"
output: html_document
---

![](C:/Users/mkhundi/Dropbox/My PC (LP256)/Documents/R/projects/miscellaneous/figures_ch3r4ds/IMG_20201112_213616.jpg){width="588"}

## 0. Welcome to the R party

![](C:/Users/mkhundi/Dropbox/My PC (LP256)/Documents/R/projects/miscellaneous/figures_ch3r4ds/IMG_20200510_070957.jpg){width="547"}

![](C:/Users/mkhundi/Dropbox/My PC (LP256)/Documents/R/projects/miscellaneous/figures_ch3r4ds/IMG_20201112_213134.jpg){width="595"}

## 1. Setup

A brief introduction to tidyverse

![](C:/Users/mkhundi/Dropbox/My PC (LP256)/Documents/R/projects/miscellaneous/figures_ch3r4ds/tidyverse_fig.png){width="149mm" height="60mm"}

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## 2. Data that will be used

The `mpg` dataframe is found in the `ggplot2` package

`mpg` contains observations collected by the US Environmental Protection Agency on 38 models of car.

Do cars with big engines use more fuel than cars with small engines?

1.  `displ`, a car's engine size, in litres.

2.  `hwy`, a car's fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance.

```{r mpg}
mpg

#check the variables in mpg
?mpg
```

## 3. Creating a ggplot graph to explore the relationship between engine size vs. fuel efficiency

Use variables `displ` on the x-axis and `hwy` on the y-axis

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

Lets interpret the output:

1.  Negative relationship

2.  Explain the ggplot functions and arguments

## 4. Turning a ggplot template into a graph

```{r, eval=FALSE}
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

```

## 5. Aesthetic mappings in ggplot2

1.  X

2.  Y

3.  Fill

4.  color

5.  shape

6.  linetype

7.  alpha

8.  

**Scaling**: ggplot2 will automatically assign a unique level of the aesthetic (here a unique color) to each unique value of the variable and a legend is added as well.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
#> Warning: Using size for a discrete variable is not advised.
```

## 6. Facets

Splitting your plot into **facets**, subplots that each display one subset of the data.

Using a single variable

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class)
```

## 7. Facet by two variables

Pay attention to the arguments of `facet_grid`

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)

```

## 8. Geometric objects in ggplot2

A **geom** is the geometrical object that a plot uses to represent data.

For example, bar charts use bar geoms, line charts use line geoms, boxplots use boxplot geoms etc

Every geom function in ggplot2 takes a `mapping` argument. However, not every aesthetic works with every geom.

```{r}
# left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# right
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

#Add the aesthetic for linetype

```

## 9. Geometric objects in ggplot2

Reproduce the graph below

![](C:/Users/mkhundi/Dropbox/My PC (LP256)/Documents/R/projects/miscellaneous/figures_ch3r4ds/geom_aesthetics.PNG)

```{r, eval=FALSE}
# make changes to reproduce the graph above
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(mapping = aes(x=  displ, y = hwy,color= drv, linetype = drv)) 
 
```

## 10. Geometric objects in ggplot2

The effect of the placement of the esthetics

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```

## **11. Statistical transformations**

Many graphs, like scatterplots, plot the raw values of your dataset. Other graphs, like bar charts, calculate new values

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

diamonds %>%
  count(cut)

?geom_bar

ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(count), group = 1))

#geom col

demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

## **12. Position adjustments**

You can colour a bar chart using either the `colour` aesthetic, or, more usefully, `fill`

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")



```

## **13. Coordinate systems**

`coord_flip()` switches the x and y axes

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()

```

```{r}
ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar()

ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y")
```

## **14. The layered grammar of graphics**

    ggplot(data = <DATA>) + 
      <GEOM_FUNCTION>(
         mapping = aes(<MAPPINGS>),
         stat = <STAT>, 
         position = <POSITION>
      ) +
      <COORDINATE_FUNCTION> +
      <FACET_FUNCTION>
