---
title: "R book club"
subtitle: "Last meeting of 2021"
date: "2021-12-03"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.width=9, fig.height=4, fig.retina=2, 
  out.width="100%",
  message = FALSE
)

# https://presentable-user2021.netlify.app/materials/act-02/03-xaringan-extras/#activity
library(flipbookr)
xaringanExtra::use_editable()
xaringanExtra::use_share_again()
```
.can-edit[
.pull-left[
### Reviewing this years progress `r emo::ji("contest")`

* We have covered chapter 1 to chapter to chapter 13 of Epi R Handbook
* Been meeting consistently (usualy more than 6 member) `r emo::ji("clap")`
* Interactive session
* New presenter **Mphatso** `r emo::ji("check")`
* New R users becoming comfortable with R
* Going through the book as a group
* Unplanned insights when going through the session
* opportunity for practising R coding and cementing R knowledge
* What else ?
]

.pull-right[
### Feedback (Areas to improve) `r emo::ji("document")`

* How would you like the sessions to be improved ?
* At times I wasn't sure on how much I should balance between beginner material and advanced 
* Should we follow the book step by step ?
* Turn around time for the videos on drop-box

]
]
---

### Next years agenda for the group `r emo::ji("key")`

* Continue with the book club?
  * Might take us up to June
  * Use 
* How do we deal with new club members who join us before finishing the book ?
* 

---
### Special tutorials focused  on particular skills `r emo::ji("key")`
* Defining custom functions in R `r emo::ji("spanner")`
* Data visualisation with ggplot2 `r emo::ji("metrics")`
* Functional programming (Understanding the purrr package)
* Introduction to gis data management in R
* Time series data analysis in R

### Other suggestions 
* Challenges session, troubleshooting session
* How to ask questions in slack
* How can we make the book club more interactive
* New users
  * Can refer them to the old videos 
* We can accommodate new members in the group



---
.panelset[

.panel[.panel-name[Code]

```{r lemur-weight, panelset=TRUE, fig.show='hide'}
library(tidyverse)
lemurs <- readRDS("lemurs.rds")

lemurs %>%
  filter(
    common_name == "Ring-Tailed Lemur",
    between(age_at_wt_y, 1, 5)
  ) %>%
  ggplot() +
  aes(x = age_at_wt_y, y = weight) +
  geom_point() +
  labs(
    x = "Age at Weight",
    y = "Weight (g)",
    title = "Weight Gain of Young Ring-Tailed Lemurs"
  )
```
]

.panel[.panel-name[Regression]

```{r echo=FALSE}
last_plot() + geom_smooth(method = "lm")
```
]

]

```{r xaringanExtra, echo=FALSE}
xaringanExtra::use_xaringan_extra(c("panelset"))
```

---
# First up: classic flipbook, using defaults to walk through code pipeline





<!-- get a flipbook version of the my_cars code chunk - pauses are where parentheses are balanced -->
---

`r chunk_reveal("my_cars")`

```{r my_cars, include = FALSE}
cars %>%
  filter(speed > 4) %>%
  ggplot() +
  aes(x = speed) + #BREAK
  aes(y = dist) + #BREAK
  geom_point(
    alpha = .8, 
    color = "blue" 
    ) + 
  aes(size = speed) #BREAK
```

---
class: inverse, center, middle


# Next up:  refer to same chunk, but use user defined breakpoints, and display output only






<!-- get another flipbook version of the my_cars code chunk, this time at the defined breaks and only showing output -->
---

`r chunk_reveal("my_cars", break_type = "user", display_type = "output", title = "### user defined break! using #BREAK message")`



---
class: inverse, center, middle

# Next up:  cycle through different lines of code





<!-- get flipbook that rotates through lines, my_fill -->
---

`r chunk_reveal("my_fill", break_type = "rotate", widths = c(1,1), title = "### cycle through lines of code marked #ROTATE")`

```{r my_fill, eval = F, echo = F}
ggplot(data = cars) +
  aes(x = speed) +
  aes(y = dist) +
  geom_point(size = 8,
             shape = 21,
             alpha = .9,
             color = "snow") +
  aes(fill = speed) +
  scale_fill_viridis_c(option = "magma") + #ROTATE
  scale_fill_viridis_c(option = "cividis") + #ROTATE
  scale_fill_viridis_c(option = "plasma") #ROTATE
```


---
class: inverse, center, middle

# Next up:  look at multiple realizations of same code





<!-- get flipbook that shows multiple realizations of the same code -->
---

`r chunk_reveal("cars_multi", break_type = 5, title = "### see multiple realizations of code")`

```{r cars_multi, include = FALSE}
cars %>% 
  sample_frac(size = 1, replace = TRUE) %>%      
  ggplot() +              
  aes(x = speed) +
  aes(y = dist) +
  geom_count(
    alpha = .7,
    color = "blue",
    size = 4
    ) + 
  geom_smooth(method = lm, se = FALSE) +
  coord_cartesian(xlim = range(cars$speed),
                  ylim = range(cars$dist)) +
  theme(legend.position = c(.9, .2))
```





