---
title: "R4Ds book club"
subtitle: "Options for variable categorisation/recoding"

author: |
  | McEwen Khundi
  |
  | London School of Hygiene & Tropical Medicine, London, UK
  | Malawi-Liverpool-Wellcome Clinical Research Programme, Blantyre, Malawi
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    df_print: paged
    theme: spacelab
    toc: true
---

<br>

## 1. Backgound

<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, message=FALSE}
library(tidyverse)    #for data manipulation
library(pmthemes)     #for ggplot themes
library(knitr)        #for tables

```

<br>

## 3. Import datasets

Import data required for the analysis.

```{r import}
library(palmerpenguins)

penguins <- palmerpenguins::penguins

```

<br>

## 4. Changing values of a factor variable

Use function in forcats i.e fct_recode

```{r tidy}
penguins <- penguins %>%
  mutate(sex2 = fct_recode(sex, "MALE"="male", "FEMALE" = "female"),
         .after = sex)
```

<br>

## 5. Changing a numerical variable into into two categories

Use if_else s

```{r}
mean(penguins$body_mass_g,na.rm =TRUE )
penguins <- penguins %>%
  mutate(above_body_mass_g = if_else(body_mass_g >= 4201.754, ">= mean mass", "< mean mass"),
         .after = body_mass_g)
```

### you can pass a function as an argument to another function

```{r}

penguins <- penguins %>%
  mutate(above_body_mass_g = if_else(body_mass_g >= mean(penguins$body_mass_g,na.rm =TRUE ), ">= mean mass", "< mean mass"),
         .after = body_mass_g)
```

<br>

## 5. Changing a variable into more than 2 categories

-   Use case_when

-   Task (Create a new categorical variable that has levels ("Mass \>=mean and Male","Mass \< mean and Male", "Female")

```{r}

```

Plot a bar graph of the var above

```{r}

```

### Other functions to consider

-   cut()

-   str_replace

-   str_sub()

<br>

## 6. Reproducibility

This reproduction of the analysis was run by:

```{r sysinfo, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

sysinfo <- Sys.info()

sysinfo <- data.frame(keyName=names(sysinfo), value=sysinfo, row.names=NULL)

sysinfo %>% kable()
```

Analysis was run at **`r Sys.time()`**, and using the following Session Info:

```{r sessioninfo, echo=FALSE, results='markdown', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```
